### 2015-03-01 paper/futurework.tex
typos, additional citation

The expression and statement layers should be extended to match the path layer.
Product types are straightforward to add to the borrow checker and will like not
pose much trouble. Sum types may be more difficult. In particular, matching by
reference allows creation of references to the payload of a sum type.
The borrow checker must ensure that the sum is not mutated while such a reference
exists. Extending the model with recursive types will require the addition of
top level (non-closure) recursive functions. Without the capability those functions
give us it is impossible to free recursive values. Those functions will necessitate
other additions, e.g. lifetime parameters. Type parameters likely do not have
significant soundness implications by themselves, but may in their interaction with
later features.

@@

Recently, Rust has added some more advanced type system features.
In particular, associated types and higher rank lifetimes.
Associated types are effectively higher kinded types.
Neither of these features will likely impact soundness, but they should still be
carefully modeled.

>>>>

The expression and statement layers should be extended to match the path layer.
Product types are straightforward to add to the borrow checker and will likely not
pose much trouble. Sum types may be more difficult. In particular, matching by
reference allows creation of references to the payload of a sum type.
The borrow checker must ensure that the sum is not mutated while such a reference
exists. 
Work on adding existential types to imperative languages can likely be adapted
for this purpose \cite{existentials}.
Extending the model with recursive types will require the addition of
top level (non-closure) recursive functions. Without the capability those functions
give us it is impossible to free recursive values. Those functions will necessitate
other additions, e.g. lifetime parameters. Type parameters likely do not have
significant soundness implications by themselves, but may in their interaction with
later features.

@@

Recently, Rust has added some more advanced type system features.
In particular, associated types and higher rank lifetime polymorphism are of interest.
Associated types are effectively functions from types to types.
Neither of these features will likely impact soundness, 
but they should still be carefully modeled.

### 2015-03-01 paper/relatedwork.tex
typos, additional citation

Cyclone later added unique pointers and a notion of borrowing. \cite{cycsafe, cycexperience}
Cyclone's version does not prevent memory leaks, which are possible both from
freeing a unique pointer that contains another live unique pointer and from
overwriting a live unique pointer. Rust statically guarantees no memory leaks in both
of these situations. Cyclone requires an explicit swap operation for using unique versions
of pointers that do not admit null values. Rust will allow any of its pointers to be null,
but guarantees that they will not be used while null. Rust also does not require explicit swaps;
Patina uses them, but that is an artifact of the model.
Cyclone considers unique pointers ``consumed'' while a borrow exists, 
whereas Rust will allow the borrowed unique pointer to be used in certain situations
(namely, reading through an immutably borrowed unique pointer is allowed).
Cyclone supports abstraction over the aliasability and kind (box versus region) of a pointer,
but Rust does not currently support any such abstraction over pointer types.

@@

Walker and Watkins discuss how regions and linear types could be combined. \cite{regionsandlinear}
Their model supports first class regions, like Cyclone, but unlike Rust.
It also introduces the idea of linear regions and provides a mechanism for
temporarily aliasing them. This could be a fruitful ground for Rust to explore
when looking to support custom allocators.
However, their model does not support mutation, which is the source of most of the complexity
of Rust's borrow checker, and this may impact its applicability to Rust.

>>>>

Cyclone later added unique pointers and a notion of borrowing \cite{cycsafe, cycexperience}.
Cyclone's version does not prevent memory leaks, which are possible both from
freeing a unique pointer that contains another live unique pointer and from
overwriting a live unique pointer. Rust statically guarantees no memory leaks in both
of these situations. Cyclone requires an explicit swap operation for using unique versions
of pointers that do not admit null values. Rust will allow any of its pointers to be null,
but guarantees that they will not be used while null. Rust also does not require explicit swaps;
Patina uses them, but that is an artifact of the model.
Cyclone considers unique pointers ``consumed'' while a borrow exists, 
whereas Rust will allow the borrowed unique pointer to be used in certain situations
(namely, reading through an immutably borrowed unique pointer is allowed).
Cyclone supports abstraction over the aliasability and kind (box versus region) of a pointer,
but Rust does not currently support any such abstraction over pointer types.

@@

Walker and Watkins discuss how regions and linear types could be combined \cite{regionsandlinear}.
Their model supports first class regions, like Cyclone, but unlike Rust.
It also introduces the idea of linear regions and provides a mechanism for
temporarily aliasing them. This could be a fruitful ground for Rust to explore
when looking to support custom allocators.
However, their model does not support mutation, which is the source of most of the complexity
of Rust's borrow checker, and this may impact its applicability to Rust.
